### 2017-11-24 thesis/contents/acknowledgements.tex

\chapter{Acknowledgements}

The journey towards finalizing this thesis has been a long one. Much has transpired in progress, and I have met many amazing people along this trek. Looking back, I would not have imagined the various opportunities, and happenstances that revealed themselves to me, and I am forever grateful.

First I thank my parents. They have sacrificed their careers and endured much hardship by immigrating to Canada, to provide me with a better future. All my achievements are only possible after I stand on their shoulders.

I met my amazing wife Jidan Zhong in 2012, at Human Brain Mapping conference in Beijing. Little did I know that I had found my kindred spirit. Without her continue to support and persistent encouragement, this thesis would not be possible.

Importantly, I thank my mentor Dr. Mojgan Hodaie. She took a chance on me when I had zero experience in neuroscience, and graciously supported me while I was having a hard time being accepted into the graduate program. I also thank her incredible patience and support for during my thesis writing. Dr. Hodaie is the most accomplished and hard-working person I have ever known. She is a world-class neurosurgeon, while at the same time balances her time between graduate mentoring, grant writing, international neurosurgical education, and at the same time multiple projects in scientific research. Every time I feel dejected from the workload, I always remind myself that Moji is doing more and that I have no excuses. It is truly my fortune to have her as my PI, and words cannot express how thankful I am.

I would like to thank my graduate committee members Dr. Karen Davis and Dr. David Mikulis. Whose extensible experience and insights were invaluable for guiding me through this process. Dr. Davis' impressive scientific acumen, and Dr. Mikulis' enthusiasm will forever stay with me.

I would also like to thank the past and present graduate and post-graduate members of the Hodaie Lab, whom I have shared an unforgettable period of my life: Danielle DeSouza, Allie Rosen, David Hayes, Brendan Behan, Matt Walker and Peter Hung. I also thank Jessica Quan, who introduced me to Dr. Hodaie and spark my journey in Neuroscience.

I was fortunate and thankful to receive support from Multiple Sclerosis Society of Canada, James F. Crother's Family Fellowship in peripheral nerve damage, and Unilever/Lipton graduate fellowships.

I thank all the research assistants over the years that helped me to perform the thankless job of download and organize patient data: Philip Yee, Udi Blankstein, Tessa Burkholder, Kevin Liang, and Erika Wharton-Shukster.

I would like to thank the friends I made over the years at the University of Toronto Chinese Martial Arts Club. We had some good times. While the club is now defunct, the friendships made there will last a lifetime.

Lastly, I would like to thank my martial arts mentor Dr. James Guo, and the Jifeng crew. The training I received and allowed me to grow beyond my limits, and instilled in me values that made this possible.