### 2017-11-24 thesis/contents/reviews/mri.tex

%\graphicspath{{D:/project/thesis_preparation/PhDtemplateLATEX_V2/ch_Background_Work/figures/}}
\section{Magnetic Resonance Imaging}
Magnetic Resonance Imaging (MRI) is a medical imaging technology that allows non-invasive visualization of the body using magnetic fields. It is a key enabler of modern healthcare and medical research, and is used in the detection and diagnosis of neuropathology.

MRI is based on the principles of nuclear magnetic resonance (NMR), where atomic nuclei, when placed in a magnetic field, can absorb and emit electromagnetic radiation. A MRI machine applies a primary magnetic field (B0) to magnetize and align the hydrogen nuclei in the body. A radio frequency (RF) field is subsequently introduced perpendicular to B0, and reorients the spin axis. This causes the nuclei to produce NMR signals detectable by the MRI scanner. Based on the precession frequency decay, the signals are recorded to construct an image of the scanned area based on the contrasts in water and fat concentrations.

\subsection{Basic Principles}
\subsubsection{Spin and Magnetic Moment}
Spin is an intrinsic property of elementary particles, and can be either positive or negative. The spin quantum number  $s$ is defined as $ s = \frac{n}{2}$, where $n \in \{0,1,2,...\}$. Therefore $s$ is a quantized value in multiples of $\frac{1}{2}$. Protons, electrons and neutrons possess spin. Nuclei with odd number of protons or neutrons have non-zero nuclear spins; while those with even number of protons and neutrons have no spin. Particles with opposite spins can pair up to produce an overall spin of zero. Hydrogen nuclei, which makes up 63$\%$ the human body, possess spins of $\frac{1}{2}$. The spin of a particle has a magnetic moment vector that behaves like a magnet with north and south poles (Figure \ref{fig:spin}).

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=3.5in]{mri/spin.pdf}
\end{tabular}
\caption{Spin of a nucleus has a magnetic moment vector $\mu$, which behaves similar to a magnet with north (N) and south (S) poles.}
\label{fig:spin}
\end{center}
\end{figure}

The spinning of nuclear particles produces angular momentum  $\textbf{J}$, that is a vector with both magnitude and direction. The largest measurable component of $J$, expressed as $I$, represents the overall quantized spin value as the spin quantum number. The corresponding quantum number is known as the magnetic quantum number $m$, and can take a total of $2I + 1$ angular momentum states.  A non-zero spin is associated with a non-zero magnetic moment ($\vec{\mu}$) via the relation $\vec{\mu} =\gamma \textbf{J}$. The z-component of the angular momentum vector (\textbf{J}) is $J_z = m \hbar$. The z-component of the magnetic moment $\mu_z$ is thus:

$$\mu_z=\gamma J_z=\gamma m \hbar$$

Where $\gamma$ is the gyromagnetic ratio, $\hbar$ is the reduced Planck constant ($=1.0545726 \times 10^{-34} m^2kg/s$ ).

\subsubsection{Effect of Magnetic Field}

In the absence of an externally applied magnetic field, the magnetic dipoles are oriented randomly. When the proton is placed in an external magnetic field $B_0$, the spin vectors become oriented with the field, either parallel or anti-parallel with respect to $B_0$ (Figure \ref{fig:externalB0}). The nuclei precess about the magnetic field direction at some frequency called Larmor frequency:

$$\omega = \gamma B$$.

Where $\gamma$ is the gyromagnetic ratio, and $B$ the magnetic field magnitude.

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=4.5in]{mri/field.pdf}
\end{tabular}
\caption{Left panel-- magnetic moments with random orientations in the absence of external magnetic field; Right panel-- magnetic moments parallel or anti-parallel aligned with the external
field} \label{fig:externalB0}
\end{center}
\end{figure}

Protons aligned in the parallel alignment are more preferred as it has a lower energy state than in anti-parallel orientation. The energy E of a magnetic moment $\vec{\mu}$ when in a magnetic field $B_0$ is given by: $E=-\vec{\mu}\cdot \textbf{B}_0=-\mu_z B_0=-\gamma m \hbar B_0$. The lower energy state of protons corresponding to m = +1/2 in a static magnetic field $B_0$ is $E(\frac{1}{2})=-\frac{1}{2}\gamma \hbar B_0$. The higher energy state of protons corresponding to m =-1/2 is given by $E(-\frac{1}{2})=\frac{1}{2}\gamma \hbar B_0$.
Consequently, the energy differences between two quantum states can be calculated as $\bigtriangleup E = \gamma \hbar B_0$ (Figure \ref{fig:energy}).

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=3in]{mri/energy}
\end{tabular}
\caption{For protons with spin quantum number of 1/2, two energy
levels and the energy difference are shown.} \label{fig:energy}
\end{center}
\end{figure}

As protons aligned in the parallel orientation are more stable, the number of spins under this state is more than that under anti-parallel orientation. This number of difference is called spin excess and the magnetic moments of the excess spins add to form the net magnetization vector $M_z$. When the spin system is at equilibrium, net magnetization lies along the z axis, parallel to $B_0$.

\subsubsection{Resonance Condition}

The steady-state of the stable spin state under $B_0 $ can be disturbed by introducing additional energy $E$ into the system.

The proton would undergo a transition from the lower energy to higher energy state by the absorption of a photon. The energy difference between the nuclear spin levels is $\bigtriangleup E = \gamma \hbar B_0$. Hence, a magnetic resonance absorption will only occur when $\nu_0 = \gamma B_0/(2 \pi)$, which is the same as Larmor frequency. This is called the \textit{resonance condition}.

\subsubsection{Relaxation}
A transverse magnetization is introduced into the system that oscillates at radio frequency (RF) in order to modify the net magnetization, this is called the RF pulse. When exposing the nuclear spin system to energy of the RF pulse, the equilibrium of the spin system is disturbed. The RF pulse is designed to introduce energy E such that the net spin direction orients at $90^\circ$. The net magnetization now is denoted as $M_{xy}$ as it aligns with the xy-plane instead of the z-axis. Individual spins now precess in phase with the RF pulse, it gives rises to the MR signal in the receiver coil.

With the termination of the RF pulse, the nuclei in the excited state will return to the equilibrium condition, this process is called \textit{relaxation}. 
The energy transmitted decays rapidly by two relaxation processes:  spin-lattice and spin-spin relaxations. These are responsible for \textit{T1 relaxation} and \textit{T2 relaxation} respectively.

\paragraph{T1 Relaxation} 
As the transverse magnetization decays, the magnetic moment gradually realigns with the z-axis of $B_0$. 
The magnitude of the magnetic moment's xy-component also gradually diminishes, and the MR signal fades. 
The $M_z$ magnetic moment recovers in proportion, and is eventually restored. 
This processed is called \textit{longitudinal relaxation} or T1 recovery, where T1 refers to the mean time for an individual nucleus to
dissipate its excess energy (heat) to the surrounding environment ( the \textit{lattice}, and therefore the processed is also named \textit{spin-lattice relaxation})and return to its equilibrium state.

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{mri/rf-pulse.pdf}
\end{tabular}
\caption{Illustration of the process of transverse relaxation. Initially, the spin precesses about the B0 field under steady-state. When a 90deg RF pulse is applied, the spin orients towards the xy-plane. When the RF pulse is switched off, the spin now relaxes towards the z-axis.  } \label{fig:energy}
\end{center}
\end{figure}

To describe how $M_z$ returns from 0 to its equilibrium during relaxation, the equation as a
function with respect to time $t$ is written as: 
$$M_z = M_0 (1 - e^{-t/T1})$$
where T1 is the time taken for approximately 63$\%$ of the longitudinal magnetization to be restored (Figure \ref{fig:T1T2}(A)). 
$M_0$ is the magnetic moment at equilibrium.

\paragraph{T2 Relaxation}

T2 relaxation is so named after the time constant of phase decoherance of the transverse magnetic moment.

Immediate after excitation, the phase precessions are synchronous, and therefore are in \textit{phase coherence}.  
However due to various reasons, the various spins begin to lose coherence, and their total sum magnetization $M_xy$ diminishes.

Phase coherence is lost in two ways:
\begin{enumerate}
\item Spins are associated with local magnetic fields, and interact randomly The energy transfer that occur in the local magnetic field interactions cause a mumulative loss of phase. It's a pure spin-spin interaction, and thereby not affected by $180^\circ$ refocusing pulse. It is also independent from the strength of $B_0$. 
The time constant which describes the return to equilibrium of the transverse magnetization, is called the \textit{spin-spin
relaxation time} or T2, and is defined in respect to time $t$ as:
$$M_{xy} =M_{xymax} e^{-t/T2}$$
where T2 is the time it takes for the transverse magnetization to decay to 37$\%$ of its original magnitude. 
T2 is always less than or equal to T1.

\item There are intrinsic inhomogeneities by the magnetic field generator, and the scanning subject. These intrinsic properties contribute to the dephasing, and has a shorter time decay than T2. This type of decay time constant is named T2*  (Figure \ref{fig:T1T2}(C)). The loss of the MR signal due to T2* is called\textit{ free induction decay} (FID). T2* is defined as follows: 
$$ \frac{1}{T_2^*} = \frac{1}{T_2} + \frac{1}{T_2inhomo}$$ 
\end{enumerate}

\begin{figure}[h]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=5.5in]{mri/T1T2}
\end{tabular}
\caption{Diagram depicting the relaxation processes of T1 (panel A), T2 (panel B) and $T_2^*$ (panel C). $T_2^*$ is shorter than T2.} \label{fig:T1T2}
\end{center}
\end{figure}

\subsubsection{MR Contrast}

MRI signal intensity depends primarily on three parameters: 
\begin{itemize}
\item Proton density: the number of excitable spins per unit volume, which determines the maximum amount of signal that can be measured from the tissue. It can be measured by minimizing T1 and T2 times. 
\item T1 time: the amount time for $M_xy$ to recover to $M_z$, to be ready for re-excitation.
\item T2 time: the amount of time for $M_xy$ signal to fade under decoherence. 
\end{itemize}
The contrast in brain tissues of the resulting MR image can be tuned by adjusting these three parameters in pulse sequence design.

\paragraph{Repetition Time}
T1 contrast is adjusted by tuning repetition time (TR), which is the length of relaxation interval between excitation pulses. When TR is long (> 1500 ms), $M_z$ recovers totally, and therefore subsequent excitation will again retain maximal signal strength; when the TR is short (< 600 ms), the $M_z$ has not totally recovered, and the next excitation pulse will not be able to gain strong MR signal. Different tissues have different intrinsic T1 differences, and therefore will show strong contrast under short TR. Therefore short TR are said to be T1 weighted.

\paragraph{Echo Time}
MR sequences have a gradient field which is used to control field  inhomogenieties and to encode spatial origin. The gradients also contribute to spin dephasing, therefore a refocusing pulse is necessary to adequately measure MR signals. The time delay between refocusing, and the measured signal under phase coherence (spin echo) is known as echo time (TE). When TE is short ($<$ 30 ms), there is little signal differences between tissue types; when TE is long ( $\geq$ 60 ms ), the intrinsic tissue differences will allow good contrast in the resulting image. Therefore long TE is said to be T2 weighted.

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{mri/spin-echo.pdf}
\end{tabular}
\caption{Illustration of the process of spin-echo. Initially, the spin precesses about the B0 field under steady-state. When a 90deg RF pulse is applied, the spins orients towards the xy-plane. When the RF pulse is switched off, the spins now relaxes towards the z-axis, and the spin undergoes phase decoherance. A 180 degrees flip angle pulse is applied to prevent spin saturation, and then a refocusing pulse is applied to bring spins back into coherance.  } \label{fig:spin-echo}
\end{center}
\end{figure}

Typically, T1-weighted spin echo (SE)  sequence is acquired with TR/TE at 340/13 ms; T2-weighted fast spin echo (FSE) at TR/TE of 3500/120 ms. 
In the case of T1-weighted MRI, images are created with short TE and short TR times. They can differentiate fat from water - with fat brighter and water darker (\url{www.mr-tip.com}) by using a gradient echo (GRE) sequence. In the brain T1-weighted scans provide good gray matter/white matter contrast. In addition, the image contrast can be increased with the use of an inversion pulse as in an MP-RAGE sequence. Due to short TR, this scan can be run very quickly, allowing the collection of high resolution 3D structural image (Figure \ref{fig:T1T2PD}, Left). These structural images can provide detailed information of the brain's shape and size, and provide the base for cortical surface reconstruction and segmentation.

The T2-weighted sequence uses a long TR and long TE. The T2-weighted sequence can be employed as a dual echo sequence. Like the T1-weighted scan, fat is
differentiated from water, but with fat dark and water bright (Figure \ref{fig:T1T2PD}, Middle).

The first or shorter echo ($T_E<30msec$) is proton density (PD) weighted or a mixture of T1 and T2. This image can provide good contrast of tissue and CSF, thus very helpful for evaluating periventricular pathology (Figure \ref{fig:T1T2PD}, Right). $T_2^*$-weighted scans use a gradient echo (GRE) sequence, with long TE and long TR. It is prone to susceptibility losses at air/tissue boundaries, but can increase contrast for certain types of tissue, such as venous blood. It is used for functional MRI detection allowing people to visualize and map the parts of the brain used to perform tasks.

\subsubsection{3D encoding}

\paragraph{Slice Selection} Recall that the resonance condition can only be satisfied when the RF pulse is equal to the Larmor frequency of the nucleus precession under the magnetic field. 
Therefore, to allow slice discrimination, the magnetic field along the z-axis is made to be inhomogeneous by introducing a linear gradient field. It's an extra coil that adds or subtracts from the main magnetic field. With the linear gradient, only a particular portion of the z-axis would share a similar unique Larmor frequency, and can be selectively excited with an RF pulse. The steepness of the gradient determines the slice thickness, the steeper the gradient, the thinner the slices.

\paragraph{Spatial Encoding}
Once the spins are excited and precess in the xy-plane, two additional gradients are applied to allow discrimination of signal measures per slice.

\begin{itemize}
\item Phase encoding: The phase encoding gradient is applied in the y-direction. It alters the Larmor frequency of the spins based the location of the y-axis, and results in phase shifts of the spins relative to each other. The phase shift is determined by the duration and amplitude of the gradient, when the gradient is switched off, all spins return to their initial rate of precession, but maintains the relative phase shift. 
\item Frequency encoding: The frequency encoding gradient is applied on the x-direction. This applies a magnetic field that increase in strength from right to left. The change in Larmor frequency alters the precession speed of the spins such that the spins towards the right precess faster than that of the left. The MR signal collected represents the entire frequency spectrum of the spins. 
\end{itemize}

The measured MR signal with both encoding methods allows recovery of spatial information from the scanned area. Application of Fourier transform will decompose the signal into component frequencies along the frequency encoding direction (x). The phase distribution along each frequencies provide information on the phase encoding direction (y). Decomposing the phase distribution requires another Fourier transform on the multiple scans of different phase gradient strengths. This is called two-dimensional Fourier transform (2D-FT), the number of phase gradient measures is called the phase encoding step.

\paragraph{K-Space}

The collected MR data are encoded in a mathematical space termed the k-space. The k-space is a two dimensional feature space, where the x-axis $k_x$ represent the measured frequency domain, and $k_y$ the phase domain. Each step in $k_y$ represents a single phase encoding step. The x-axis origin $k_x=0$ is the gradient iso-center, and the frequency distribution is symmetric about the x-origin. The 2D-FT operation is applied to the k-space to transform the frequency domain information into image.

\begin{figure}[htb]
\begin{center}
%\begin{tabular}{c}
\includegraphics[width=5in]{mri/T1T2PD}
%\end{tabular}
\caption{Brain scan using T1 (left), T2 (middle) and proton density (right) measurements.}
\label{fig:T1T2PD}
\end{center}
%\vspace{-0.15in}

\end{figure}