### 2018-02-28 contents/glossary.tex
Updates from ShareLaTeX

\newacronym{tn}{TN}{Trigeminal Neuralgia}
\newacronym{dwi}{DWI}{Diffusion Weighted Imaging}
\newacronym{dti}{DTI}{Diffusion Tensor Imaging}
\newacronym{hardi}{HARDI}{High angular resolution diffusion imaging}
\newacronym{mri}{MRI}{Magnetic Resonance Imaging}
\newacronym{fmri}{fMRI}{Functional Magnetic Resonaonce Imaging}
\newacronym{dsi}{DSI}{Diffusion Spectrum Imaging}
\newacronym{sh}{SH}{Spherical Harmonics}
\newacronym{fod}{FOD}{Fiber orientation distribution}
\newacronym{csd}{CSD}{Constrained spherical deconvolution}
\newacronym{fa}{FA}{Fractional Anisotropy}
\newacronym{rd}{RD}{Radial Anisotropy}
\newacronym{ad}{AD}{Axial Anisotropy}
\newacronym{md}{MD}{Mean Anisotropy}
\newacronym{adc}{ADC}{Apperant diffusion coefficient}
\newacronym{ms}{MS}{Multiple Sclerosis}
\newacronym{mstn}{MS-TN}{Symptomatic Trigeminal Neuralgia Secondary To Multiple Sclerosis}
\newacronym{trg}{TRG}{Trigeminal Root Ganglion}
\newacronym{gp}{GP}{Gaussian Process}
\newacronym{svm}{SVM}{Support Vector Machine}
\newacronym{cnv}{CN V}{Cranial Nerve 5, the trigeminal nerve}
\newacronym{dmri}{dMRI}{Diffusion Magnetic Resonance Imaging}
\newacronym{nvc}{NVC}{Neurovascular Compression}
\newacronym{rez}{REZ}{Root Entry Zone}
\newacronym{tpt}{TPT}{Trigeminopontothalamic tract}
\newacronym{s1}{S1}{primary sensory cortex}
\newacronym{s2}{S2}{secondary sensory cortex}
\newacronym{acc}{ACC}{anterior cingulate cortex}
\newacronym{ofc}{OFC}{orbitofrontal cortex}
\newacronym{pfc}{PFC}{prefrontal cortex}
\newacronym{nac}{NAc}{nucleus accumbens}
\newacronym{bold}{BOLD}{Blood-Oxygen-Level Dependent}
\newacronym{vmpo}{VMpo}{ventralmedial posterior thalamus}
\newacronym{te}{TE}{Echo Time}
\newacronym{tr}{TR}{Repetition Time}
\newacronym{gre}{GRE}{Gradient Echo}
\newacronym{dnn}{DNN}{Deep neural network}
\newacronym{eeg}{EEG}{Electroencephalography}
\newacronym{ml}{ML}{Machine Learning}
\newacronym{pca}{PCA}{Principal components analysis}
\newacronym{epi}{EPI}{Echo Planar Imaging}
\newacronym{ap}{AP}{Anisotropic Power Image}
\newacronym{sagit}{SAGIT}{Selective automated group integrated tractography}
\newacronym{nos}{NOS}{Normalized overlap score}
\newacronym{xst}{XST}{eXtended Streamline Tractography}
\newacronym{flair}{FLAIR}{Fluid-attenuated inversion recovery imaging}
\newacronym{mprage}{MP-RAGE}{Magnetization-prepared 180 degrees radio-frequency pulses and rapid gradient-echo imaging}
\newacronym{mdwi}{MDWI}{Mean diffusion weighted image}
\newacronym{cc}{CC}{Cross correlation}
\newacronym{MI}{MI}{Mutual information}

\glsaddall