### 2018-07-23 contents/contributions.tex

\chapter{Statement of Contributions}
% \addcontentsline{toc}{chapter}{Statement of Contributions}

{\parskip=12pt
\noindent
Dr. David Qixiang Chen (author) - all aspect of the work, including some healthy subject enrollment and data collection, experiment design, data analysis, methods development, data interpretation, and thesis preparation.

\noindent
Dr. Mojgan Hodaie (supervisor) - mentorship, assistance with experiment design, assistance with data interpretation, and assistance with thesis preparation.

\noindent
Dr. Karen D. Davis (committee member) - mentorship, assistance with thesis preparation.

\noindent
Dr. David Mikulis (commitee member) - mentorship, assistance with thesis preparation.

\noindent
Dr. Jidan Zhong - assistance with experiment design, assistance with data interpretation, assistance with method development

\noindent
Dr. Dave J. Hayes - assistance with experiment design, assistance with data interpretation

\noindent
Powell Chu - assistance with data collection, assistance with method development
}